### 2024-10-31 Relatorio/paper.tex
Clean Repository
Cleaned repository and code to prepare for fresh start.

Class diagram complete, code can now begin. This commit includes the client class, the validator class and enumerators.

% This template has been tested with LLNCS DOCUMENT CLASS -- version 2.21 (12-Jan-2022)

% !TeX spellcheck = pt_PT-Portuguese
% !LTeX: language= pt_PT-Portuguese
% !TeX encoding = utf8
% !TeX program = pdflatex
% !BIB program = bibtex
% -*- coding:utf-8 mod:LaTeX -*-

% "a4paper" enables:
%
%  - easy print out on DIN A4 paper size
%
% One can configure a4 vs. letter in the LaTeX installation. So it is configuration dependend, what the paper size will be.
% This option  present, because the current word template offered by Springer is DIN A4.
% We accept that DIN A4 cause WTFs at persons not used to A4 in USA.
%
% "runningheads" enables:
%
%  - page number on page 2 onwards
%  - title/authors on even/odd pages
%
% This is good for other readers to enable proper archiving among other papers and pointing to
% content. Even if the title page states the title, when printed and stored in a folder, when
% blindly opening the folder, one could hit not the title page, but an arbitrary page. Therefore,
% it is good to have title printed on the pages, too.
%
% To disable outputting page headers and footers, remove "runningheads"
\documentclass[runningheads,a4paper,english]{llncs}[2022/01/12]

% backticks (`) are rendered as such in verbatim environments.
% See following links for details:
%   - https://tex.stackexchange.com/a/341057/9075
%   - https://tex.stackexchange.com/a/47451/9075
%   - https://tex.stackexchange.com/a/166791/9075
\usepackage{upquote}

% Used to add images
\usepackage{graphicx}

% Set English as language and allow to write hyphenated"=words
%
% Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
% english has to go last to set it as default language
\usepackage[ngerman,main=english]{babel}
%
% Hint by http://tex.stackexchange.com/a/321066/9075 -> enable "= as dashes
\addto\extrasenglish{\languageshorthands{ngerman}\useshorthands{"}}
%
% Fix by https://tex.stackexchange.com/a/441701/9075
\usepackage{regexpatch}
\makeatletter
\edef\switcht@albion{%
  \relax\unexpanded\expandafter{\switcht@albion}%
}
\xpatchcmd*{\switcht@albion}{ \def}{\def}{}{}
\xpatchcmd{\switcht@albion}{\relax}{}{}{}
\edef\switcht@deutsch{%
  \relax\unexpanded\expandafter{\switcht@deutsch}%
}
\xpatchcmd*{\switcht@deutsch}{ \def}{\def}{}{}
\xpatchcmd{\switcht@deutsch}{\relax}{}{}{}
\edef\switcht@francais{%
  \relax\unexpanded\expandafter{\switcht@francais}%
}
\xpatchcmd*{\switcht@francais}{ \def}{\def}{}{}
\xpatchcmd{\switcht@francais}{\relax}{}{}{}
\makeatother

% Links behave as they should. Enables "\url{...}" for URL typesettings.
% Allow URL breaks also at a hyphen, even though it might be confusing: Is the "-" part of the address or just a hyphen?
% See https://tex.stackexchange.com/a/3034/9075.
\usepackage[hyphens]{url}

% When activated, use text font as url font, not the monospaced one.
% For all options see https://tex.stackexchange.com/a/261435/9075.
% \urlstyle{same}

% Improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother

% nicer // - solution by http://tex.stackexchange.com/a/98470/9075
% DO NOT ACTIVATE -> prevents line breaks
%\makeatletter
%\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
%\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
%\makeatother

% This is the modern package for "Computer Modern".
% In case this gets activated, one has to switch from cmap package to glyphtounicode (in the case of pdflatex)
\usepackage[%
    rm={oldstyle=false,proportional=true},%
    sf={oldstyle=false,proportional=true},%
    % By using 'variable=true' the monospaced font can be used as variable font (with differents widths per letter)
    % However, this makes listings look ugly.
    tt={oldstyle=false,proportional=true,variable=false},%
    qt=false%
]{cfr-lm}

% Has to be loaded AFTER any font packages. See https://tex.stackexchange.com/a/2869/9075.
\usepackage[T1]{fontenc}

% Character protrusion and font expansion. See http://www.ctan.org/tex-archive/macros/latex/contrib/microtype/

\usepackage[
  babel=true, % Enable language-specific kerning. Take language-settings from the languge of the current document (see Section 6 of microtype.pdf)
  expansion=alltext,
  protrusion=alltext-nott, % Ensure that at listings, there is no change at the margin of the listing
  final % Always enable microtype, even if in draft mode. This helps finding bad boxes quickly.
        % In the standard configuration, this template is always in the final mode, so this option only makes a difference if "pros" use the draft mode
]{microtype}

% \texttt{test -- test} keeps the "--" as "--" (and does not convert it to an en dash)
\DisableLigatures{encoding = T1, family = tt* }

%\DeclareMicrotypeSet*[tracking]{my}{ font = */*/*/sc/* }%
%\SetTracking{ encoding = *, shape = sc }{ 45 }
% Source: http://homepage.ruhr-uni-bochum.de/Georg.Verweyen/pakete.html
% Deactiviated, because does not look good

\usepackage{graphicx}

% Diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
% Slashbox is not available in texlive (due to licensing) and also gives bad results. Thus, we use diagbox
\usepackage{diagbox}

\usepackage{xcolor}

% Code Listings
\usepackage{listings}

\definecolor{eclipseStrings}{RGB}{42,0.0,255}
\definecolor{eclipseKeywords}{RGB}{127,0,85}
\colorlet{numb}{magenta!60!black}

% JSON definition
% Source: https://tex.stackexchange.com/a/433961/9075

\lstdefinelanguage{json}{
    basicstyle=\normalfont\ttfamily,
    commentstyle=\color{eclipseStrings}, % style of comment
    stringstyle=\color{eclipseKeywords}, % style of strings
    numbers=left,
    numberstyle=\scriptsize,
    stepnumber=1,
    numbersep=8pt,
    showstringspaces=false,
    breaklines=true,
    frame=lines,
    % backgroundcolor=\color{gray}, %only if you like
    string=[s]{"}{"},
    comment=[l]{:\ "},
    morecomment=[l]{:"},
    literate=
        *{0}{{{\color{numb}0}}}{1}
         {1}{{{\color{numb}1}}}{1}
         {2}{{{\color{numb}2}}}{1}
         {3}{{{\color{numb}3}}}{1}
         {4}{{{\color{numb}4}}}{1}
         {5}{{{\color{numb}5}}}{1}
         {6}{{{\color{numb}6}}}{1}
         {7}{{{\color{numb}7}}}{1}
         {8}{{{\color{numb}8}}}{1}
         {9}{{{\color{numb}9}}}{1}
}

\lstset{
  % everything between (* *) is a latex command
  escapeinside={(*}{*)},
  %
  language=json,
  %
  showstringspaces=false,
  %
  extendedchars=true,
  %
  basicstyle=\footnotesize\ttfamily,
  %
  commentstyle=\slshape,
  %
  % default: \rmfamily
  stringstyle=\ttfamily,
  %
  breaklines=true,
  %
  breakatwhitespace=true,
  %
  % alternative: fixed
  columns=flexible,
  %
  numbers=left,
  %
  numberstyle=\tiny,
  %
  basewidth=.5em,
  %
  xleftmargin=.5cm,
  %
  % aboveskip=0mm,
  %
  % belowskip=0mm,
  %
  captionpos=b
}

% Enable Umlauts when using \lstinputputlisting.
% See https://stackoverflow.com/a/29260603/873282 für details.
% listingsutf8 did not work in June 2020.
\lstset{literate=
  {á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'i}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
  {Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
  {à}{{\`a}}1 {è}{{\`e}}1 {ì}{{\`i}}1 {ò}{{\`o}}1 {ù}{{\`u}}1
  {À}{{\`A}}1 {È}{{\'E}}1 {Ì}{{\`I}}1 {Ò}{{\`O}}1 {Ù}{{\`U}}1
  {ä}{{\"a}}1 {ë}{{\"e}}1 {ï}{{\"i}}1 {ö}{{\"o}}1 {ü}{{\"u}}1
  {Ä}{{\"A}}1 {Ë}{{\"E}}1 {Ï}{{\"I}}1 {Ö}{{\"O}}1 {Ü}{{\"U}}1
  {â}{{\^a}}1 {ê}{{\^e}}1 {î}{{\^i}}1 {ô}{{\^o}}1 {û}{{\^u}}1
  {Â}{{\^A}}1 {Ê}{{\^E}}1 {Î}{{\^I}}1 {Ô}{{\^O}}1 {Û}{{\^U}}1
  {Ã}{{\~A}}1 {ã}{{\~a}}1 {Õ}{{\~O}}1 {õ}{{\~o}}1
  {œ}{{\oe}}1 {Œ}{{\OE}}1 {æ}{{\ae}}1 {Æ}{{\AE}}1 {ß}{{\ss}}1
  {ű}{{\H{u}}}1 {Ű}{{\H{U}}}1 {ő}{{\H{o}}}1 {Ő}{{\H{O}}}1
  {ç}{{\c c}}1 {Ç}{{\c C}}1 {ø}{{\o}}1 {å}{{\r a}}1 {Å}{{\r A}}1
}

% For easy quotations: \enquote{text}
% This package is very smart when nesting is applied, otherwise textcmds (see below) provides a shorter command
\usepackage[autostyle=true]{csquotes}

% Enable using "`quote"' - see https://tex.stackexchange.com/a/150954/9075
\defineshorthand{"`}{\openautoquote}
\defineshorthand{"'}{\closeautoquote}

% Nicer tables (\toprule, \midrule, \bottomrule)
\usepackage{booktabs}

% Extended enumerate, such as \begin{compactenum}
\usepackage{paralist}

% Bibliopgraphy enhancements
%  - enable \cite[prenote][]{ref}
%  - enable \cite{ref1,ref2}
% Alternative: \usepackage{cite}, which enables \cite{ref1, ref2} only (otherwise: Error message: "White space in argument")

% Doc: http://texdoc.net/natbib
\usepackage[%
  square,        % for square brackets
  comma,         % use commas as separators
  numbers,       % for numerical citations;
  %sort           % orders multiple citations into the sequence in which they appear in the list of references;
  sort&compress  % as sort but in addition multiple numerical citations
                 % are compressed if possible (as 3-6, 15);
]{natbib}

% In the bibliography, references have to be formatted as 1., 2., ... not [1], [2], ...
\renewcommand{\bibnumfmt}[1]{#1.}

% Enable hyperlinked author names in the case of \citet
% Source: https://tex.stackexchange.com/a/76075/9075
\usepackage{etoolbox}
\makeatletter
\patchcmd{\NAT@test}{\else \NAT@nm}{\else \NAT@hyper@{\NAT@nm}}{}{}
\makeatother

% Prepare more space-saving rendering of the bibliography
% Source: https://tex.stackexchange.com/a/280936/9075
\SetExpansion
[ context = sloppy,
  stretch = 30,
  shrink = 60,
  step = 5 ]
{ encoding = {OT1,T1,TS1} }
{ }

% Put figures aside a text
\usepackage[rflt]{floatflt}

% Enable nice comments
\usepackage{pdfcomment}

\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}

% Compatibality with packages todo, easy-todo, todonotes
\newcommand{\todo}[1]{\commentatside{#1}}

% Compatiblity with package fixmetodonotes
\newcommand{\TODO}[1]{\commentatside{#1}}

% Put footnotes below floats
% Source: https://tex.stackexchange.com/a/32993/9075
\usepackage{stfloats}
\fnbelowfloat

\usepackage[group-minimum-digits=4,per-mode=fraction]{siunitx}

% Enable that parameters of \cref{}, \ref{}, \cite{}, ... are linked so that a reader can click on the number an jump to the target in the document
\usepackage{hyperref}

% Enable hyperref without colors and without bookmarks
\hypersetup{
  hidelinks,
  colorlinks=true,
  allcolors=black,
  pdfstartview=Fit,
  breaklinks=true
}

% Enable correct jumping to figures when referencing
\usepackage[all]{hypcap}

\usepackage[caption=false,font=footnotesize]{subfig}

\usepackage{mindflow}

% Extensions for references inside the document (\cref{fig:sample}, ...)
% Enable usage \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link
% That means, "Figure 5" is a full link instead of just "5".
\usepackage[capitalise,nameinlink]{cleveref}

\crefname{section}{Sec.}{Sec.}
\Crefname{section}{Section}{Sections}
\crefname{listing}{List.}{List.}
\crefname{listing}{Listing}{Listings}
\Crefname{listing}{Listing}{Listings}
\crefname{lstlisting}{Listing}{Listings}
\Crefname{lstlisting}{Listing}{Listings}

\usepackage{lipsum}

% For demonstration purposes only
% These packages can be removed when all examples have been deleted
\usepackage[math]{blindtext}
\usepackage{mwe}
\usepackage[realmainfile]{currfile}
\usepackage{tcolorbox}
\tcbuselibrary{listings}

%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
\DeclareFontFamily{U}{MnSymbolC}{}
\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
\DeclareFontShape{U}{MnSymbolC}{m}{n}{
  <-6>    MnSymbolC5
  <6-7>   MnSymbolC6
  <7-8>   MnSymbolC7
  <8-9>   MnSymbolC8
  <9-10>  MnSymbolC9
  <10-12> MnSymbolC10
  <12->   MnSymbolC12%
}{}
\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}

\usepackage{xspace}
\newcommand{\eg}{e.g.,\ }
\newcommand{\ie}{i.e.,\ }

% Enable hyphenation at other places as the dash.
% Example: applicaiton\hydash specific
\makeatletter
\newcommand{\hydash}{\penalty\@M-\hskip\z@skip}
% Definition of "= taken from http://mirror.ctan.org/macros/latex/contrib/babel-contrib/german/ngermanb.dtx
\makeatother

% Add manual adapted hyphenation of English words
% See https://ctan.org/pkg/hyphenex and https://tex.stackexchange.com/a/22892/9075 for details
% Does not work on MiKTeX, therefore disabled - issue reported at https://github.com/MiKTeX/miktex-packaging/issues/271
% \input{ushyphex}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

% Add copyright
%
% This is recommended if you intend to send the version to colleagues
% See https://ctan.org/pkg/llncsconf for details
\iffalse
  % state: intended | submitted | llncs
  % you can add "crop" if the paper should be cropped to the format Springer is publishing
  \usepackage[intended]{llncsconf}

  \conference{name of the conference}

  % in case of "proceedings" (final version!)
  % example: \llncs{Anonymous et al. (eds). \emph{Proceedings of the International Conference on \LaTeX-Hacks}, LNCS~42. Some Publisher, 2016.}{0042}
  % 0042 denotes an example start page
  \llncs{book editors and title}{0042}
\fi

% Enable copy and paste of text from the PDF
% Only required for pdflatex. It "just works" in the case of lualatex.
% Alternative: cmap or mmap package
% mmap enables mathematical symbols, but does not work with the newtx font set
% See: https://tex.stackexchange.com/a/64457/9075
% Other solutions outlined at http://goemonx.blogspot.de/2012/01/pdflatex-ligaturen-und-copynpaste.html and http://tex.stackexchange.com/questions/4397/make-ligatures-in-linux-libertine-copyable-and-searchable
% Trouble shooting outlined at https://tex.stackexchange.com/a/100618/9075
%
% According to https://tex.stackexchange.com/q/451235/9075 this is the way to go
\input glyphtounicode
\pdfgentounicode=1

\begin{document}

\title{Gestão de Alojamentos Turísticos}
% If Title is too long, use \titlerunning
%\titlerunning{Short Title}
\subtitle{Programação Orientada a Objetos}

% Single insitute
\author{Enrique George Rodrigues Nº 28602}

% If there are too many authors, use \authorrunning
%\authorrunning{First Author et al.}

\institute{Instituto Politécnico do Cávado e do Ave\\
	Licenciatura Engenharia de Sistemas Informáticos\\
	\vspace{6pt}
	16 de Outubro de 2024}

%% Multiple insitutes - ALTERNATIVE to the above
% \author{%
%     Firstname Lastname\inst{1} \and
%     Firstname Lastname\inst{2}
% }
%
%If there are too many authors, use \authorrunning
%  \authorrunning{First Author et al.}
%
%  \institute{
%      Insitute 1\\
%      \email{...}\and
%      Insitute 2\\
%      \email{...}
%}

\maketitle

\renewcommand{\abstractname}{Resumo.} % Change the abstract title to "Resumo"
\begin{abstract}

Este projeto tem como objetivo o desenvolvimento de uma aplicação para a gestão de alojamentos turísticos, permitindo a administração eficiente de registos, reservas e consultas relacionadas com clientes e alojamentos. A solução foi implementada em C\#, utilizando o paradigma de programação orientada a objetos, de forma a garantir uma estrutura modular e escalável. As principais funcionalidades incluem o registo de clientes, a gestão de reservas, o processamento de check-ins e check-outs, bem como o controlo de pagamentos. A aplicação destaca-se pelo uso de dicionários como estruturas de dados eficientes, permitindo um acesso rápido e otimizado a registos de clientes, reservas e alojamentos, com uma complexidade temporal de $O(1)$ para operações de consulta e inserção. A persistência de dados é gerida em formato JSON, proporcionando uma solução leve e prática para operadores turísticos.

\renewcommand{\abstractname}{Resumo}

% Redefine keywords manually since llncs doesn't define \keywordsname
\def\keywords{\vspace{.5em}
	{\textbf{Palavras-chave:}}\,\relax%
}

\keywords{registos, consultas, reservas, check-in, clientes, alojamentos}
\end{abstract}

\section{Introdução}
\label{sec:introduction}

O setor do turismo enfrenta desafios crescentes na gestão eficiente de alojamentos, exigindo soluções que garantam agilidade, precisão e uma experiência satisfatória para os clientes. Este projeto visa desenvolver um sistema de gestão de alojamentos turísticos que permita a administração eficaz de registos de clientes, reservas e alojamentos. A implementação foi realizada em C\#, uma linguagem reconhecida pela sua robustez e flexibilidade, permitindo o desenvolvimento de aplicações escaláveis e eficientes.

Para garantir um desempenho otimizado, o sistema utiliza dicionários como estruturas de dados principais, que se baseiam em tabelas de hash, proporcionando acesso rápido e eficiente a informações cruciais, com uma complexidade temporal de $O(1)$ para operações de consulta e inserção. Além disso, a persistência de dados é gerida em formato JSON, facilitando a integração e manipulação das informações. Embora a eficiência e a funcionalidade sejam os pilares deste projeto, a aplicação também foi concebida com atenção às boas práticas de programação, promovendo um código limpo e manutenível. Desta forma, o sistema não apenas atende às necessidades atuais do setor, mas também está preparado para futuras evoluções e inovações na gestão de alojamentos turísticos.

Este documento descreve o trabalho prático da unidade curricular de Programação Orientada a Objetos, parte integrante da Licenciatura em Engenharia de Sistemas Informáticos no Instituto Politécnico do Cávado e do Ave.

O enunciado deste trabalho prático pode ser encontrado no anexo titulado ``Trabalho\_POO\_ESI\_2024\_2025``. O código fonte está disponível no Github, veja \href[]{LINK}.

\section{Padrões e Práticas de Programação}
\label{sec:codingstyle}

Para garantir a consistência, legibilidade e manutenção da base de código, este projeto segue os padrões de programação estabelecidos. Estes padrões estão alinhados com o Guia de Estilo da Microsoft para C\# \cite{microsoftcsharpstyle}, promovendo uniformidade em todo o projeto.

\subsection{Convenções de Nomeação}

A adoção de uma convenção de nomeação clara melhora a clareza do código. As seguintes convenções foram utilizadas ao longo do projeto:

\begin{itemize}
	\item \textbf{Classes e Structs}: \texttt{ClassName}
	\item \textbf{Métodos}: \texttt{MethodName()}
	\item \textbf{Propriedades}: \texttt{PropertyName}
	\item \textbf{Variáveis de Instância e Locais}: \texttt{variableName}
	\item \textbf{Constantes}: \texttt{CONSTANT\_NAME}
	\item \textbf{Parâmetros}: \texttt{parameterName}
	\item \textbf{Enums}: \texttt{EnumName}
\end{itemize}

Evita-se o uso de abreviações excessivas e nomes confusos, promovendo a clareza no código.

\subsection{Regras de Formatação}

A formatação consistente é fundamental para a legibilidade:

\begin{itemize}
	\item \textbf{Indentação}: 4 espaços por nível de indentação, evitando tabs.
	\item \textbf{Comprimento da Linha}: Linhas limitadas a um máximo de 120 caracteres.
	\item \textbf{Chaves}: As chaves de abertura são colocadas na próxima linha da declaração de controle.
	\item \textbf{Espaçamento}: Espaços em branco são usados para separar operadores e ao redor de chaves.
	\item \textbf{Comentários}: Comentários descritivos são utilizados para explicar seções complexas do código, seguindo um estilo consistente.
\end{itemize}

\subsection{Práticas de Programação}

Seguir as melhores práticas garante a qualidade do código:

\begin{itemize}
	\item Utilizam-se nomes claros para variáveis e métodos, tornando o código autoexplicativo.
	\item Escreve-se código claro e conciso, evitando complexidade desnecessária.
	\item Segue-se o Guia de Estilo da Microsoft para C\# \cite{microsoftcsharpstyle} para formatação e estilo gerais, garantindo a conformidade com os padrões da indústria.
\end{itemize}

\subsection{Controlo de Versões}

Utiliza-se Git para o controle de versões, facilitando a gestão eficaz da base de código:

\begin{itemize}
	\item Cria-se uma nova branch para cada funcionalidade ou correção de erro.
	\item As branches são criadas a partir da branch principal para desenvolvimento, com nomes descritivos que refletem a funcionalidade ou erro abordado.
	\item Evita-se fazer commits diretamente na branch principal, garantindo a estabilidade.
	\item As branches de funcionalidade são regularmente fundidas de volta na branch principal após testes completos.
\end{itemize}

\subsection{Documentação}

Utiliza-se \textit{XML Documentation Comments} para documentar o código, permitindo a geração automática de documentação da API:

\begin{itemize}
	\item Segue-se a sintaxe do XML para comentários, documentando métodos, propriedades e classes.
	\item A documentação inclui descrições breves, descrições de parâmetros, descrições de valores de retorno e exemplos de uso.
	\item Utilizam-se tags como \texttt{<summary>}, \texttt{<param>}, \texttt{<returns>} e \texttt{<example>} para estruturar corretamente os comentários.
	\item Define-se o autor nos comentários de cabeçalho de classes para novos ficheiros ou seções significativas de código.
\end{itemize}

\section{Estrutura de Classes}
\label{sec:classes}

O diagrama de classes apresentado descreve a estrutura lógica do sistema de gestão de reservas para alojamentos turísticos, com uma arquitetura desenhada para centralizar e organizar informações essenciais sobre clientes, reservas, alojamentos e pagamentos. Este modelo, por além de otimizar o fluxo de trabalho, promove a extensibilidade e modularidade, facilitando tanto as expansões futuras como a manutenção do sistema.

No diagrama, cada classe representa uma componente fundamental do sistema. A classe \textit{Client} (cliente) armazena informações sobre cada utilizador, incluindo dados pessoais como nome, e-mail e endereço. A classe \textit{Accommodation} (alojamento) representa as propriedades ou quartos disponíveis para reserva, contendo informações como o tipo de alojamento, capacidade e preço por noite. A classe \textit{Reservation} (reserva) liga os clientes aos alojamentos, registando os detalhes da estadia, incluindo as datas de check-in e check-out, o custo total e o estado do pagamento.

Para processar pagamentos, a classe \textit{Payment} (pagamento) regista os dados de cada transação associada a uma reserva, como o montante pago, a data de pagamento e o método utilizado. A classe \textit{BookingManager} (gestor de reservas) atua como o controlador principal do sistema, fornecendo métodos para adicionar e remover clientes, criar e cancelar reservas, gerir alojamentos e processar pagamentos.

O sistema introduz a interface \textit{ManageableEntity}, implementada pelas classes \textit{Clients}, \textit{Reservations} e \textit{Accommodations}. Esta interface unifica os métodos básicos de gestão, como \textit{Add()}, \textit{Remove()}, \textit{Import()} e \textit{Export()}, permitindo que o \textit{BookingManager} interaja com estas classes de forma uniforme. Desta forma, o \textit{BookingManager} atua como a face do sistema, permitindo direcionar chamadas para a classe apropriada com base no tipo de entidade a ser manipulada.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\textwidth]{Images/POO_Class_Diagram.png}
	\caption{Diagrama de Classes do Sistema de Gestão de Alojamentos Turísticos}
	\label{fig:diagrama_classes}
\end{figure}

\section{Conclusão}
\label{sec:conclusion}

w

\subsubsection*{Agradecimentos}

Gostaria de expressar o meu sincero agradecimento ao Professor Luís Ferreira, pela sua orientação, apoio e \textit{feedback} valiosos ao longo do desenvolvimento deste projeto. A sua ajuda foi fundamental no desenvolvimento desta solução informática.

%%% ===============================================================================
%%% Bibliography
%%% ===============================================================================

\renewcommand{\bibsection}{\section*{Referências}} % requried for natbib to have "References" printed and as section*, not chapter*
% Use natbib compatbile splncs04nat style.
% It does provide all features of splncs04.bst, but is developed in a clean way.
% Source: https://github.com/tpavlic/splncs04nat
\bibliographystyle{splncs04nat}
\begingroup
  \microtypecontext{expansion=sloppy}
  \small % ensure correct font size for the bibliography
  \bibliography{paper}
\endgroup

% Enfore empty line after bibliography
\ \\
%
Todas as ligações foram consultadas pela última vez no dia 25 de outubro de 2024.

%%% ===============================================================================
%\appendix
%\addcontentsline{toc}{chapter}{APPENDICES}

%\listoffigures
%\listoftables
%%% ===============================================================================

%%% ===============================================================================
%\section{My first appendix}\label{sec:appendix1}
%%% ===============================================================================
\end{document}